---
title: "Commercial Value"
author: "Min-Yang Lee"
date: "`r format(Sys.time(), '%B %d, %Y')`" 
output:
  html_document:
    df_print: paged
    keep_md: yes
    fig_caption: yes

---

```{r setup, include=FALSE}
###############################################################################
# Purpose: 	Assemble gross commercial revenue by stock and year. Aggregates CAMS
#           landings and value from the species/statistical-area level up to the
#           stock level, folds in non-Northeast-managed species, and writes the
#           commercial deliverable plus the stock keyfile.
#
# Inputs:
#  - data_folder/main/species_area_landings_{vintage}.Rds   (fmp_value_datapull.R)
#  - data_folder/main/stock_area_definitions_{vintage}.Rds  (fmp_value_datapull.R)
#  - data_folder/main/fmp_listing_{vintage}.Rds             (fmp_value_datapull.R)
#  - data_folder/main/Other_species_landings_{vintage}.Rds  (fmp_value_datapull.R)
#  - data_folder/raw/deflators_{vintage}.Rds                (FRED_extraction.R; unused, see Notes)
#
# Outputs:
#  - data_folder/main/stock_keyfile_{vintage}.Rds
#  - data_folder/main/Commercial_Value_{vintage}.Rds
#  - data_folder/main/Commercial_Value_{vintage}.csv
#  - the knit HTML document and, via keep_md, its markdown
#
# Execution order
# Stage 2 of the production pipeline, and the first of the two documents.
# Sources FRED_extraction.R and fmp_value_datapull.R from the `extract_data` chunk
# below. Nothing calls this document -- it is knit by hand.
#
# THIS DOCUMENT MUST BE KNIT BEFORE writing/Recreational_Value.Rmd. It is the sole
# producer of stock_keyfile_{vintage}.Rds, which that document reads. Knitting them
# in the other order on a clean data folder fails with a missing-file error whose
# cause is not obvious.
#
# Notes
# Commercial value is reported in NOMINAL dollars, by design. The deflators are
# read in below and the `process_deflators` chunk is deliberately dormant; neither
# is an oversight or unfinished work.
#
# The extraction scripts are sourced from a chunk marked eval=FALSE. That is a
# runtime convenience, not a sign the step is optional -- pulling from CAMS is slow,
# so the extract is run once and the document is then knit against the saved
# vintage as many times as needed. See REPO-STATUS.md.
###############################################################################

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(comment = FALSE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(here)
library(tidyverse)
library(data.table)
library(reshape2)
library(ggpubr)
library(kableExtra)
library(glue)
library(survey)
library(fredr)
library(conflicted)
conflicts_prefer(dplyr::filter)
conflicts_prefer(lubridate::year)
conflicts_prefer(dplyr::summarise)
conflicts_prefer(dplyr::arrange)
options(scipen=999)
options(survey.lonely.psu = "certainty")

here::i_am("writing/Commercial_Value.Rmd")


```

```{r extract_data, eval=FALSE, include=TRUE}

#Some of the data prep takes a long time. 
#Instead of pulling data every time I want to compile this doc, I have split off the code to do so into the following pieces.

# Get GDP implicit price deflators
source(here("R_code", "data_extraction_processing", "extraction","FRED_extraction.R"))
# Get commercial data
source(here("R_code", "data_extraction_processing", "extraction","fmp_value_datapull.R"))

```

```{r readin_data, include=FALSE}
# There is no vintage global. Each extraction script stamps its output with the
# date it ran, and the vintage for the whole pipeline is recovered here by globbing
# for the commercial landings file and taking the newest date string. Everything
# else read below is looked up with that same string, which is why the extraction
# scripts all have to be run on the same calendar day.
vintage_string<-list.files(here("data_folder","main"), pattern=glob2rx("species_area_landings_*Rds"))
vintage_string<-gsub("species_area_landings_","",vintage_string)
vintage_string<-gsub(".Rds","",vintage_string)
vintage_string<-max(vintage_string)

CAMS_revenue<-readRDS(file=here("data_folder", "main", glue("species_area_landings_{vintage_string}.Rds")))
stock_area_definitions<-readRDS(file=here("data_folder", "main", glue("stock_area_definitions_{vintage_string}.Rds")))
stock_area_definitions<-stock_area_definitions %>%
  rename(stockarea=area_name2)
fmp_listing<-readRDS(file=here("data_folder", "main", glue("fmp_listing_{vintage_string}.Rds")))

#landings and value for non northeast managed species.
non_NEM_landings<-readRDS(file=here("data_folder","main",glue("Other_species_landings_{vintage_string}.Rds")))



# Read in Deflators
# Read but never used: commercial value is reported nominal by design. Note this is
# the one input that lives in data_folder/raw rather than main, and it is looked up
# with the vintage discovered from the commercial landings file above, not from its
# own filename -- so FRED_extraction.R has to have run the same day.
deflators<-readRDS(file=here("data_folder","raw",glue("deflators_{vintage_string}.Rds")))
```

```{r process_deflators, include=FALSE, eval=FALSE}
# Dormant by design -- commercial value is not deflated. Unlike the extract_data
# chunk above, this one is not merely skipped for speed: it is retained as a record
# of how deflation would be applied if it were ever wanted. As written it would not
# run, because params$deflate_year has no corresponding params: block in the YAML.

deflate_base<-deflators %>%
  filter(year==params$deflate_year)%>%
  pull(value)

# multiply a NOMINAL value by the deflation multiplier to get REAL 'base-year' dollars
deflators<-deflators %>%
  mutate(deflation_multiplier  = deflate_base/value) %>%
  select(year,series_id,deflation_multiplier)

```

# Introduction

This document describes how we assemble Gross Revenues for Commercial.

# Gross Revenues for Commercial

Gross revenues aggregated to the species and statistical area were extracted from CAMS.  We aggregated to the stockarea using a modification of the ``cams_garfo.cfg_statarea_stock`` table.  This does not have all stock definitions in it, fortunately most of the missing ones are Unit stocks.  We used the ``STOCKEFF`` website to look up which species are unit stocks and which have areas.  For stocks not in ``STOCKEFF,`` we used the stock assessment and relevant fishery management documents.

```{r dataclean_commercial}

# Pull in stock area definitions.  Drop out landings from Canada or the 700s.

# The join happens on area as a character string, matching how both tables carry it
# out of Oracle; the numeric coercion is deferred until after the join so the
# statistical-area filter below can be written as numbers.
CAMS_revenue<-CAMS_revenue %>%
  left_join(stock_area_definitions, by=join_by(itis_tsn==itis_tsn, area==area))%>%
  mutate(area=as.numeric(area))%>%
  filter(area %in% c(465,464,
                     511,512,513,515,521,522,525,526,
                     533,534,537,538,539,541,542,543,561,562,
                     611,612,613,614,615,616,621,622,623,624,
                     625,626,627,628,629,
                     631,632,633,634,
                     635,636,637,638,639)) %>%
  mutate(missing=is.na(stockarea))
# There is still a little slop, A bit of skates that do not match. 
#table(CAMS_revenue$itis_name, CAMS_revenue$missing)

# Some do not have a definition. Classify those as unit stocks
# Two spellings arrive here: unmatched rows get "UNIT" from the line below, while
# rows that matched carry the "Unit" written by the case_when blocks in
# fmp_value_datapull.R. The second mutate collapses them so the group_by that
# follows does not split one stock into two.
CAMS_revenue<-CAMS_revenue %>%
  mutate(stockarea=if_else(is.na(stockarea),"UNIT",stockarea))%>%
  mutate(stockarea=if_else(stockarea=="Unit","UNIT",stockarea)) %>%
  select(-c(common_name, area_name))

#make sure stockarea is all filled in.
missing_check <-CAMS_revenue %>%
  filter(is.na(stockarea))
stopifnot(nrow(missing_check)==0)

# Aggregate landings and value to the stock-year level.
CAMS_revenue<-CAMS_revenue%>%
  group_by(council, itis_name, itis_tsn, stockarea, itis_sci_name, fmp, year) %>%
  summarise(commercial_value=sum(value),
            commercial_landed_pounds=sum(landings)) %>%
  ungroup()


# Species with no council have no stock-area definitions to allocate against, so
# they come out of Oracle aggregated to year only and are carried as unit stocks.
non_NEM_landings<-non_NEM_landings %>%
  select(council, itis_name, itis_tsn, itis_sci_name, fmp, year, value, landings) %>%
    group_by(council, itis_name, itis_tsn, itis_sci_name, fmp, year) %>%
  summarise(commercial_value=sum(value),
            commercial_landed_pounds=sum(landings)) %>%
  ungroup() %>%
  mutate(stockarea="UNIT")

CAMS_revenue<-rbind(CAMS_revenue,non_NEM_landings)
CAMS_revenue<-CAMS_revenue %>%
  arrange(council, itis_name, itis_tsn, itis_sci_name, fmp, year)


```

## Which stocks do I have

Print out a table of stocks by manager. This is just an easy way to make sure we have what we are expecting.  The stocks info was extracted from CAMS, so it uses the current stock boundaries. 


```{r keyfiles}
stock_keyfile<-CAMS_revenue %>%
  select(c(council, fmp, itis_tsn, itis_name, itis_sci_name, stockarea)) %>%
  group_by(council, fmp, itis_tsn, itis_name, itis_sci_name, stockarea) %>% 
    slice(1) %>%
  ungroup() %>%
  arrange(council, fmp, itis_tsn, itis_name, itis_sci_name, stockarea)

# writing/Recreational_Value.Rmd reads this file. It is the only reason the two
# documents have to be knit in order, and nothing enforces that ordering.
saveRDS(stock_keyfile, file=here("data_folder","main",glue("stock_keyfile_{vintage_string}.Rds")))



ASFMC<-stock_keyfile %>%
  filter(council == "ASMFC" | council=="ASMFC/NEFMC") %>% 
  arrange(fmp, itis_name)


MAFMC<-stock_keyfile %>%
  filter(council == "MAFMC" | council=="MAFMC/NEFMC")  %>% 
  arrange(fmp, itis_name)

NEFMC<-stock_keyfile %>%
  filter(council == "NEFMC" | council=="MAFMC/NEFMC"  | council=="ASMFC/NEFMC") %>% 
  arrange(fmp, itis_name)

Non_Managed<-stock_keyfile %>%
  filter(is.na(council)) %>% 
  arrange(fmp, itis_name)

out_of_geog<-fmp_listing %>%
  filter(fmp  %in% c("SERO FMP","Highly Migratory Species"))  %>% 
  arrange(fmp, itis_name)
```

### ASMFC stocks

```{r ASFMC_stocks}
 knitr::kable(ASFMC %>% select(-c(council, itis_tsn)), caption='ASMFC stocks',format.args = list(big.mark = ","), digits=1, align=c("l", "l","l","r") ) 

```
### MAFMC stocks

```{r MAFMC_stocks}
 knitr::kable(MAFMC  %>% select(-c(council, itis_tsn)), caption='MAFMC stocks',format.args = list(big.mark = ","), digits=1, align=c("l", "l","l","r") ) 

```

### NEFMC stocks


```{r NEFMC_stocks}

 knitr::kable(NEFMC  %>% select(-c(council, itis_tsn)), caption='NEFMC stocks',format.args = list(big.mark = ","), digits=1, align=c("l", "l","l","r") ) 

```


### Non-Northeast

These are a mixed of state managed stocks, HMS, and SERO managed stocks.


```{r Non-Managed}
 knitr::kable(Non_Managed  %>% select(-c(council, itis_tsn,stockarea)), caption='Non-Northest Region stocks',format.args = list(big.mark = ","), digits=1, align=c("l", "l","l","r") ) 

```

## Format the commercial data.

```{r format_and_export}
Commercial_Value<-CAMS_revenue %>%
  select(-commercial_landed_pounds) %>%
  pivot_wider(names_from=year, values_from=commercial_value) %>%
  relocate(council, fmp, itis_tsn, itis_sci_name, itis_name, stockarea)

saveRDS(Commercial_Value, file=here("data_folder","main",glue("Commercial_Value_{vintage_string}.Rds")))

write.csv(x = Commercial_Value, file=here("data_folder","main",glue("Commercial_Value_{vintage_string}.csv")))

```




